### 2016-07-15 testbed_http.tex
Almost there, just missing the write image part

In Section~\ref{sec:overlay_fs}, the boot loader was configured to call
the initial \ac{RAM} filesystem. Instead of that, now we present how the
the initial \ac{RAM} filesystem in the boot partition is configured to
download and cache the root filesystem from a \ac{HTTP} server before
mounting it. The initial \ac{RAM} filesystem is initialized by the
\texttt{\$\{HTTPROOTDIR\}/boot/init.gz} script. This is first called
when the system is booting up before mounting the root filesystem and
parses the arguments specified in \texttt{\$\{HTTPROOTDIR\}/boot/cmdline.txt}.
We modify the arguments in this file to download, cache and mount the
\ac{HTTP} root image filesystem (i.e. the \texttt{httproot.img} file) with a
customized script.

@@

Make the script executable
\begin{lstlisting}[]
# chmod +x ${HTTPROOTDIR}/etc/initramfs-tools/scripts/http
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

The variable (boot=http) that we specified in \texttt{cmdline.txt} tells the
\texttt{init} script to call the script above rather than the default scripts
that mounts the root filesystem from a block device or a \ac{NFS} shared
folder. The \texttt{init} script will call the function \texttt{http\_mount\_root}
which in this case can be consired as the \texttt{main} function.

As in any other shell scripts, this means that code in the global namespace
will be executed by the interpretor before the function \texttt{http\_mount\_root}
is entered.
%Due to the variable that we specified in \texttt{cmdline.txt},
Hence, line 9 will be the first command executed. This line creates a directory
in \texttt{initramfs} named
\texttt{imagedir} that we will use to hold the root image.

Next, line 12 to line 44 will parse the additional arguments that we provided in
\texttt{cmdline.txt}.

The \ac{Raspi}s does not have enough \ac{RAM} to hold entire root
filesystem, so we instead download it to the partition that, as specified in
\texttt{fstab} later, will later be mounted as the \texttt{HOME} directory
when the system starts. This will enable the \ac{Raspi}s to cache the root image
and use the previously downloaded root image again after reboots (see the
function \texttt{http\_fetch\_image} starting from line 56)\fxnote{is there a better way to reference to a function within the script? This reference will be invalid if the script is slightly modified}.

Edit fstab to look like this (exclude the last line if a persistent
home storage directory is not desired):
\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$\{HTTPROOTDIR\}/etc/fstab}@>
<@\textcolor{gray}{
---------------------------------------------------------------}
\Reactivatenumber @>
proc            /proc           proc    defaults          0       0
/dev/mmcblk0p1  /boot           vfat    defaults          0       2
/dev/loop0      /               ext4    defaults,noatime  0       1
/dev/mmcblk0p2  /home           ext4    defaults,noatime  0       0
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Chroot into the image
% CHROOT to OS image
\begin{lstlisting}[]
$ cd $HTTPROOTDIR
# mount -t proc proc proc/
# mount --bind /sys sys/
# mount --bind /dev dev/
# mount --bind /dev/pts dev/pts
# proot -q qemu-arm-static -S ${HTTPROOTDIR}
# export PS1="(chroot) $PS1"
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Create new initramfs (remember that your kernel might be different)
\begin{lstlisting}[]
(chroot) # mkinitramfs -o /boot/init.gz -k 4.4.13+
(chroot) # mkinitramfs -o /boot/init-v7.gz -k 4.4.13-v7+
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Exit chroot
\begin{lstlisting}[]
(chroot) # exit
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Unmount:
\begin{lstlisting}[]
$ cd ..
# umount ${HTTPROOTIMAGE}/{dev/pts,proc,sys,dev}
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

\subsection{Prepare sd-card}

@@

% Define variable
% export IMAGE="2016-05-27-raspbian-jessie-lite"
% export WORKDIR="${HOME}/Raspbian"
\begin{lstlisting}[]
$ export HTTPDEV="/dev/mmcblk0"
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

@@

\subsection{Compressing root image (Optional)}

PROBABLY A BETTER IDEA TO USE BTRFS TRANSPARENT COMPRESSION

\begin{lstlisting}[]
# apt-get install squashfs-tools
# mount -o loop ${HTTPROOTIMAGE}.img ${HTTPROOTDIR}
# mksquashfs ${HTTPROOTDIR} compressed_${HTTPROOTIMAGE}.img
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

\subsection{Try it out}

Upload the root image (\$\{HTTPROOTIMAGE\}.img) to the HTTP server used in the
configurations above. Then, unplug the memory card from the computer.
Insert it into a \ac{Raspi} and
turn it on. It is recommended that the \ac{Raspi} is turned on with a monitor
connected the first time the newly created Linux installation is tested to
assure that Raspbian starts up correctly.

PRESENT FILES ON HTTP. E.G. HTTPROOT.IMG AND IMAGES.CSV

\begin{itemize}
    \item Copy http root image (\$\{HTTPROOTIMAGE\}.img) a HTTP server
    \item Create images.csv
\end{itemize}

@@

As of this writing, there is unfortunately a bug in wget that makes wget not showing
its progress bar while downloading the root image at startup. Instead, it may appear
to be unable to connect although it is downloading, so keep patient.

\begin{lstlisting}[]
Connecting to <HTTP SERVER IP> (<HTTP SERVER IP>:<PORT>)
image.csv       100% |*******************************************************| 1287   0:00:00 ETA
Connecting to <HTTP SERVER IP> (<HTTP SERVER IP>:<PORT>)
_
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

REMEMBER TO COPY-PASTE ALL CONTENT FROM SCRIPT INTO LATEX AGAIN IN CASE SOME
SCRIPT CHANGED, BUT WAS NOT UPDATED IN LATEX

>>>>

\subsection{HTTP Image Fetch Script}
In Section~\ref{sec:overlay_fs}, the boot loader was configured to call
the initial \ac{RAM} filesystem. Instead of that, now we present how the
the initial \ac{RAM} filesystem in the boot partition is configured to
download and cache the root filesystem from a \ac{HTTP} server before
mounting it. The initial \ac{RAM} filesystem is initialized by the
\texttt{\$\{HTTPROOTDIR\}/boot/init.gz} script. This is first called
when the system is booting up before mounting the root filesystem and
parses the arguments specified in \texttt{\$\{HTTPROOTDIR\}/boot/cmdline.txt}.
We modify the arguments in this file to download, cache and mount the
\ac{HTTP} root image filesystem (i.e. the \texttt{httproot.img} file) with a
customized script.

@@

Once the script has been saved, we make the script executable by doing:
\begin{lstlisting}[]
# chmod +x ${HTTPROOTDIR}/etc/initramfs-tools/scripts/http
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

When using \texttt{boot=http} in the \texttt{cmdline.txt} file, this
indicates the \texttt{initramfs} initialization script to call the script
above instead of other default scripts that mount the root filesystem.
Once called, the initialization script will first execute the
\texttt{mount\_root()} function which can be consired as the \texttt{main}
function for this script. As in any other shell scripts, this means
that any code in the global namespace will executed by the interpretor
before the function \texttt{mount\_root()} is entered when booting. Hence,
line 9 of the \texttt{http} script will be the first command executed.
This line creates a directory in the \texttt{initramfs} named
\texttt{imagedir} that we will use to hold the root image. Next, lines 12
to 44 will parse the additional arguments that we provided in
the \texttt{cmdline.txt} file.

Since the \ac{Raspi}s do not have enough \ac{RAM} to hold the entire
root filesystem, we download it to the partition that, as specified
in \texttt{\$\{HTTPROOTDIR\}/etc/fstab} file later, will be mounted as the
\texttt{\$HOME} directory when the system starts. This will enable the
\ac{Raspi}s to cache the root image and use the previously downloaded
root image again after any reboots. For details, see the function
\texttt{http\_fetch\_image()} just after the variables verification.
We edit the mentioned \texttt{fstab} file with root privileges to look
as the following (exclude the last line if a persistent home storage
directory is not desired):

% \fxnote{is there a better way to reference to a function within the script? This reference will be invalid if the script is slightly modified}.
\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$\{HTTPROOTDIR\}/etc/fstab}@>
<@\textcolor{gray}{
-----------------------------------------------}
\Reactivatenumber @>
proc            /proc           proc    defaults          0       0
/dev/mmcblk0p1  /boot           vfat    defaults          0       2
/dev/loop0      /               ext4    defaults,noatime  0       1
/dev/mmcblk0p2  /home           ext4    defaults,noatime  0       0
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

\subsection{Initial RAM Filesystem for HTTP Fetch}
We enter again in \texttt{chroot} mode into the image to make the new
\texttt{initramfs} files. We note that for other testbed administrator
kernels, the files might be different. The commands for these tasks are:

% CHROOT to HTTP image
\begin{lstlisting}[]
$ cd $HTTPROOTDIR
# mount -t proc proc proc/
# mount --bind /sys sys/
# mount --bind /dev dev/
# mount --bind /dev/pts dev/pts
# chroot ${HTTPROOTDIR} /usr/bin/qemu-arm-static /bin/bash
# export PS1="(chroot) $PS1"
(chroot) # mkinitramfs -o /boot/init.gz -k 4.4.13+
(chroot) # mkinitramfs -o /boot/init-v7.gz -k 4.4.13-v7+
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

In case the \texttt{chroot} command does not work, it is possible to use
the alternative \texttt{proot} command as described in
Section~\ref{sec:chroot}. If successfully executed, the boot partition for
the SD card is ready to be written. We exit the \texttt{chroot} environment
and unmount all the partitions as follows:

\begin{lstlisting}[]
(chroot) # exit
$ cd ..
# umount ${HTTPROOTIMAGE}/{dev/pts,proc,sys,dev}
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

\subsection{Writing SD Cards for HTTP Fetch}

@@

\begin{lstlisting}[]
$ export HTTPDEV="/dev/mmcblk0"
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

@@

% IT WILL BE GOOD TO TRY THE COMPRESSED IMAGE BUT MAYBE LATER:

% \subsection{Compressing root image (Optional)}
% PROBABLY A BETTER IDEA TO USE BTRFS TRANSPARENT COMPRESSION
% \begin{lstlisting}[]
% # apt-get install squashfs-tools
% # mount -o loop ${HTTPROOTIMAGE}.img ${HTTPROOTDIR}
% # mksquashfs ${HTTPROOTDIR} compressed_${HTTPROOTIMAGE}.img
% \end{lstlisting}
% \FloatBarrier
% \vspace{-5mm}

\subsection{HTTP Images File and Setup Test}

We upload the \texttt{\$\{HTTPROOTIMAGE\}.img} file to the \ac{HTTP}
server used in the configurations above. Finally, in the \ac{HTTP} server,
the testbed administrator has to create
\texttt{http://<SERVERIP>/.../images.csv} file. This file contains the
information regarding which image should be download for each \ac{Raspi}. A
sample file is the following:

@@

Once this file is set in the \ac{HTTP} server, the testbed
administrator needs to unplug the memory card from its \ac{PC} and insert
it into a \ac{Raspi} and turn it on. For this part, it is recommended that the
\ac{Raspi} is turned on with a monitor connected to assure that Raspbian
starts up correctly. At the time of writing this manuscript, a bug in
the \texttt{wget} command does not permit to show the image download progress
bar at startup. Instead, it may appear to be unable to connect.
Nevertheless, the image should be downloading. On average this process takes
XX minutes. If all the procedure has been successfully executed,
the testbed administrator should observe the \ac{Raspi} properly booting up
after the downloading the root image file. Finally, the testbed adminstrator
only needs to copy one time the boot partition to all the cards and
boot the remaining \ac{Raspi}s. This concludes the procedure.

% \begin{lstlisting}[]
% Connecting to <HTTP SERVER IP> (<HTTP SERVER IP>:<PORT>)
% image.csv       100% |*******************************************************| 1287   0:00:00 ETA
% Connecting to <HTTP SERVER IP> (<HTTP SERVER IP>:<PORT>)
% \end{lstlisting}
% \FloatBarrier
% \vspace{-5mm}

% DO WE REALLY NEED TO DO THIS? -->
% REMEMBER TO COPY-PASTE ALL CONTENT FROM SCRIPT INTO LATEX AGAIN IN CASE SOME
% SCRIPT CHANGED, BUT WAS NOT UPDATED IN LATEX
